---
title: "Assignment"
output: html_document
---

## Loading and preprocessing the data

```{r, echo=TRUE}
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?


total number of steps taken per day

```{r, echo=TRUE}
totalnumberofsteps <- aggregate(steps ~ date, data,  sum)
hist(totalnumberofsteps$steps , xlab="total number of steps", ylab="day")
```

```{r meanandmedian, echo=TRUE}
meanoftotalnumberofsteps <- mean(totalnumberofsteps$steps)
medianoftotalnumberofsteps <- median(totalnumberofsteps$steps)
```

Total number of steps taken per day

- Mean `r as.integer(meanoftotalnumberofsteps)` 

- Median `r medianoftotalnumberofsteps` 

##What is the average daily activity pattern?

```{r averagegraph, echo=TRUE, fig.height=4}
averagesteps <- aggregate(steps  ~ interval , data,  mean)

plot(averagesteps$interval, averagesteps$steps,  type = "l", xlab="interval", ylab="average numbers of taken")
```



```{r maxnumberofinterval, echo=TRUE, fig.height=4}
#max number of interval

maxstepsInterval <- data$interval[which.max(data$steps)]
```

steps that contains max number of steps is `r maxstepsInterval` 

##Imputing missing values

```{r nacounts, echo=TRUE}
nacount <- sum(is.na(data$steps))
```

There are `r nacount` NA values in the data

Filling NA's with interval means
```{r imputing, echo=TRUE, fig.height=4}
library(plyr)
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
data <- ddply(data, ~ interval, transform, steps = impute.mean(steps))
totalnumberofsteps <- aggregate(steps ~ date, data,  sum)
hist(totalnumberofsteps$steps , xlab="total number of steps", ylab="day")

```

As we can see increase with center values(10000-15000 steps day ) and decrease in extreme values(center of bar the bar grap gets bigger and others gets smaller)

##Are there differences in activity patterns between weekdays and weekends?
```{r weekdays, echo=TRUE, fig.height=20}

library(lattice)

data$date <- as.Date(data$date)

weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')

data$date <-  factor((weekdays(data$date) %in% weekdays1)+1L,
                    levels=1:2, labels=c('weekend', 'weekday'))


par(mfrow=c(2,1)) 

plot <- xyplot(steps ~ interval | date, xlab = "interval", ylab = "Number of Steps", data=data, type = "l",main = "Average Steps by Interval", layout = c(1, 2))


print(plot, position = c(0, .3, 1, .9), more = TRUE)
```

 
